Component response rate variation drives stability in large complex systems

### 2018-07-29 notebook/SI.tex
Discussion draft written and SI edits

Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
matrices found is greater than 1 (compare column 5 to column 3), and
this ratio increases with increasing \(S\) (column 1). Hence, more
randomly created complex systems (\(\mathbf{M}\)) are stable given
variation in \(\gamma\) than when \(\gamma = 1\). Note that feasibility
results were ommited for the table above, but are
\protect\hyperlink{Feasibility}{reported below}.

@@

While the foundational work of
May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
to complex networks, much attention has been given specifically to
ecological networks of interacting species. In these networks, the
matrix \(\mathbf{M}\) is interpreted as a community matrix and each row
and column is interpreted as a single species. The effect that the
density of any species \(i\) has on the population dynamics of species
\(j\) is found in \(M_{ij}\), meaning that \(\mathbf{M}\) holds the
effects of pair-wise interactions between \(S\)
species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
While May's original
work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
randomly assembled communities, recent work has specifically looked at
more restricted ecological communities including competitive networks
(all off-diagonal elements of \(\mathbf{M}\) are negative), mutualist
networks (all off-diagonal elements of \(\mathbf{M}\) are positive), and
predator-prey networks (for any pair of \(i\) and \(j\), the effect of
\(i\) on \(j\) is negative and \(j\) on \(i\) is positive, or vice
versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
In general, competitor and mutualist networks tend to be unstable, while
predator-prey networks tend to be highly stabilising.

@@

\begin{longtable}[]{@{}llllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}llllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}rrrrrr@{}}

@@

A1\_stabilised\tabularnewline

@@

2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline

@@

\hypertarget{connectance}{\section{Sensitivity of inter-connectivity (C)
values}\label{connectance}}

>>>>

Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
matrices found is greater than 1 whenever \(S > 10\) (compare column 5
to column 3), and this ratio increases with increasing \(S\) (column 1).
Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
given variation in \(\gamma\) than when \(\gamma = 1\). Note that
feasibility results were omitted for the table above, but are
\protect\hyperlink{Feasibility}{reported below}.

@@

While the foundational work of
May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
to complex networks, much attention has been given specifically to
ecological networks of interacting species. In these networks, the
matrix \(\mathbf{A}\) is interpreted as a community matrix and each row
and column is interpreted as a single species. The effect that the
density of any species \(i\) has on the population dynamics of species
\(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds the
effects of pair-wise interactions between \(S\)
species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
While May's original
work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
randomly assembled communities, recent work has specifically looked at
more restricted ecological communities including competitive networks
(all off-diagonal elements of \(\mathbf{A}\) are negative), mutualist
networks (all off-diagonal elements of \(\mathbf{A}\) are positive), and
predator-prey networks (for any pair of \(i\) and \(j\), the effect of
\(i\) on \(j\) is negative and \(j\) on \(i\) is positive, or vice
versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
In general, competitor and mutualist networks tend to be unstable, while
predator-prey networks tend to be highly stabilising.

@@

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
3 & 229 & 99771 & 231 & 99769 & 0 & 2\tabularnewline
4 & 701 & 99299 & 704 & 99296 & 0 & 3\tabularnewline
5 & 1579 & 98421 & 1587 & 98413 & 0 & 8\tabularnewline
6 & 3218 & 96782 & 3253 & 96747 & 6 & 41\tabularnewline
7 & 5519 & 94481 & 5619 & 94381 & 23 & 123\tabularnewline
8 & 9062 & 90938 & 9237 & 90763 & 77 & 252\tabularnewline
9 & 13436 & 86564 & 13729 & 86271 & 230 & 523\tabularnewline
10 & 18911 & 81089 & 19303 & 80697 & 505 & 897\tabularnewline
11 & 25594 & 74406 & 25961 & 74039 & 1011 & 1378\tabularnewline
12 & 33207 & 66793 & 33382 & 66618 & 1724 & 1899\tabularnewline
13 & 41160 & 58840 & 41089 & 58911 & 2655 & 2584\tabularnewline
14 & 50575 & 49425 & 49894 & 50106 & 3777 & 3096\tabularnewline
15 & 59250 & 40750 & 57892 & 42108 & 4824 & 3466\tabularnewline
16 & 67811 & 32189 & 65740 & 34260 & 5634 & 3563\tabularnewline
17 & 75483 & 24517 & 73056 & 26944 & 5943 & 3516\tabularnewline
18 & 82551 & 17449 & 79878 & 20122 & 5780 & 3107\tabularnewline
19 & 88030 & 11970 & 85204 & 14796 & 5417 & 2591\tabularnewline
20 & 92254 & 7746 & 89766 & 10234 & 4544 & 2056\tabularnewline
21 & 95233 & 4767 & 93002 & 6998 & 3695 & 1464\tabularnewline
22 & 97317 & 2683 & 95451 & 4549 & 2803 & 937\tabularnewline
23 & 98508 & 1492 & 97122 & 2878 & 1991 & 605\tabularnewline
24 & 99240 & 760 & 98407 & 1593 & 1216 & 383\tabularnewline
25 & 99669 & 331 & 99082 & 918 & 739 & 152\tabularnewline
26 & 99871 & 129 & 99490 & 510 & 452 & 71\tabularnewline
27 & 99938 & 62 & 99732 & 268 & 240 & 34\tabularnewline
28 & 99985 & 15 & 99888 & 112 & 108 & 11\tabularnewline
29 & 99990 & 10 & 99951 & 49 & 46 & 7\tabularnewline
30 & 100000 & 0 & 99981 & 19 & 19 & 0\tabularnewline
31 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
32 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
33 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
34 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 56 & 99944 & 56 & 99944 & 0 & 0\tabularnewline
3 & 3301 & 96699 & 3301 & 96699 & 0 & 0\tabularnewline
4 & 34446 & 65554 & 34446 & 65554 & 0 & 0\tabularnewline
5 & 86520 & 13480 & 86520 & 13480 & 0 & 0\tabularnewline
6 & 99683 & 317 & 99683 & 317 & 0 & 0\tabularnewline
7 & 99998 & 2 & 99998 & 2 & 0 & 0\tabularnewline
8 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
9 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
10 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
11 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
12 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}rrrrrrr@{}}

@@

A1\_stabilised & A1\_destabilised\tabularnewline

@@

2 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
3 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
4 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
5 & 1 & 99999 & 1 & 99999 & 0 & 0\tabularnewline
6 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
7 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
8 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
9 & 20 & 99980 & 21 & 99979 & 0 & 1\tabularnewline
10 & 20 & 99980 & 22 & 99978 & 0 & 2\tabularnewline
11 & 38 & 99962 & 39 & 99961 & 0 & 1\tabularnewline
12 & 64 & 99936 & 66 & 99934 & 0 & 2\tabularnewline
13 & 87 & 99913 & 91 & 99909 & 0 & 4\tabularnewline
14 & 157 & 99843 & 159 & 99841 & 0 & 2\tabularnewline
15 & 215 & 99785 & 227 & 99773 & 0 & 12\tabularnewline
16 & 293 & 99707 & 310 & 99690 & 0 & 17\tabularnewline
17 & 383 & 99617 & 408 & 99592 & 0 & 25\tabularnewline
18 & 443 & 99557 & 473 & 99527 & 3 & 33\tabularnewline
19 & 642 & 99358 & 675 & 99325 & 4 & 37\tabularnewline
20 & 836 & 99164 & 887 & 99113 & 7 & 58\tabularnewline
21 & 1006 & 98994 & 1058 & 98942 & 10 & 62\tabularnewline
22 & 1153 & 98847 & 1228 & 98772 & 20 & 95\tabularnewline
23 & 1501 & 98499 & 1593 & 98407 & 30 & 122\tabularnewline
24 & 1841 & 98159 & 1996 & 98004 & 40 & 195\tabularnewline
25 & 2146 & 97854 & 2316 & 97684 & 58 & 228\tabularnewline
26 & 2643 & 97357 & 2809 & 97191 & 119 & 285\tabularnewline
27 & 3034 & 96966 & 3258 & 96742 & 158 & 382\tabularnewline
28 & 3690 & 96310 & 3928 & 96072 & 201 & 439\tabularnewline
29 & 4257 & 95743 & 4532 & 95468 & 290 & 565\tabularnewline
30 & 4964 & 95036 & 5221 & 94779 & 424 & 681\tabularnewline
31 & 5627 & 94373 & 5978 & 94022 & 452 & 803\tabularnewline
32 & 6543 & 93457 & 6891 & 93109 & 666 & 1014\tabularnewline
33 & 7425 & 92575 & 7777 & 92223 & 818 & 1170\tabularnewline
34 & 8540 & 91460 & 8841 & 91159 & 1071 & 1372\tabularnewline
35 & 9526 & 90474 & 9842 & 90158 & 1337 & 1653\tabularnewline
36 & 10617 & 89383 & 10891 & 89109 & 1624 & 1898\tabularnewline
37 & 12344 & 87656 & 12508 & 87492 & 2021 & 2185\tabularnewline
38 & 13675 & 86325 & 13877 & 86123 & 2442 & 2644\tabularnewline
39 & 15264 & 84736 & 15349 & 84651 & 2870 & 2955\tabularnewline
40 & 17026 & 82974 & 17053 & 82947 & 3363 & 3390\tabularnewline
41 & 18768 & 81232 & 18614 & 81386 & 3905 & 3751\tabularnewline
42 & 20791 & 79209 & 20470 & 79530 & 4579 & 4258\tabularnewline
43 & 23150 & 76850 & 22754 & 77246 & 5217 & 4821\tabularnewline
44 & 25449 & 74551 & 24184 & 75816 & 6285 & 5020\tabularnewline
45 & 27702 & 72298 & 26464 & 73536 & 6754 & 5516\tabularnewline
46 & 30525 & 69475 & 28966 & 71034 & 7646 & 6087\tabularnewline
47 & 32832 & 67168 & 31125 & 68875 & 8487 & 6780\tabularnewline
48 & 36152 & 63848 & 33865 & 66135 & 9479 & 7192\tabularnewline
49 & 38714 & 61286 & 36242 & 63758 & 10125 & 7653\tabularnewline
50 & 41628 & 58372 & 38508 & 61492 & 11036 & 7916\tabularnewline
51 & 44483 & 55517 & 41023 & 58977 & 11704 & 8244\tabularnewline
52 & 48134 & 51866 & 44287 & 55713 & 12573 & 8726\tabularnewline
53 & 51138 & 48862 & 46721 & 53279 & 13223 & 8806\tabularnewline
54 & 54261 & 45739 & 49559 & 50441 & 13757 & 9055\tabularnewline
55 & 57647 & 42353 & 52403 & 47597 & 14324 & 9080\tabularnewline
56 & 60630 & 39370 & 55293 & 44707 & 14669 & 9332\tabularnewline
57 & 63647 & 36353 & 57787 & 42213 & 15103 & 9243\tabularnewline
58 & 66961 & 33039 & 60439 & 39561 & 15450 & 8928\tabularnewline
59 & 69968 & 30032 & 63708 & 36292 & 15246 & 8986\tabularnewline
60 & 72838 & 27162 & 66270 & 33730 & 15177 & 8609\tabularnewline
61 & 75609 & 24391 & 68873 & 31127 & 15006 & 8270\tabularnewline
62 & 77999 & 22001 & 71318 & 28682 & 14538 & 7857\tabularnewline
63 & 80616 & 19384 & 73517 & 26483 & 14510 & 7411\tabularnewline
64 & 83089 & 16911 & 76209 & 23791 & 13784 & 6904\tabularnewline
65 & 85150 & 14850 & 78086 & 21914 & 13412 & 6348\tabularnewline
66 & 86908 & 13092 & 80437 & 19563 & 12477 & 6006\tabularnewline
67 & 88671 & 11329 & 82379 & 17621 & 11718 & 5426\tabularnewline
68 & 90537 & 9463 & 84483 & 15517 & 10878 & 4824\tabularnewline
69 & 91969 & 8031 & 86233 & 13767 & 10033 & 4297\tabularnewline
70 & 93181 & 6819 & 87914 & 12086 & 9070 & 3803\tabularnewline
71 & 94330 & 5670 & 89200 & 10800 & 8401 & 3271\tabularnewline
72 & 95324 & 4676 & 90833 & 9167 & 7359 & 2868\tabularnewline
73 & 96143 & 3857 & 91805 & 8195 & 6726 & 2388\tabularnewline
74 & 96959 & 3041 & 93065 & 6935 & 5900 & 2006\tabularnewline
75 & 97543 & 2457 & 93987 & 6013 & 5222 & 1666\tabularnewline
76 & 97969 & 2031 & 94900 & 5100 & 4481 & 1412\tabularnewline
77 & 98497 & 1503 & 95756 & 4244 & 3809 & 1068\tabularnewline
78 & 98744 & 1256 & 96442 & 3558 & 3269 & 967\tabularnewline
79 & 99045 & 955 & 96942 & 3058 & 2837 & 734\tabularnewline
80 & 99276 & 724 & 97528 & 2472 & 2329 & 581\tabularnewline
81 & 99481 & 519 & 97996 & 2004 & 1894 & 409\tabularnewline
82 & 99556 & 444 & 98321 & 1679 & 1597 & 362\tabularnewline
83 & 99691 & 309 & 98722 & 1278 & 1227 & 258\tabularnewline
84 & 99752 & 248 & 98943 & 1057 & 1015 & 206\tabularnewline
85 & 99833 & 167 & 99144 & 856 & 837 & 148\tabularnewline
86 & 99895 & 105 & 99346 & 654 & 642 & 93\tabularnewline
87 & 99925 & 75 & 99461 & 539 & 530 & 66\tabularnewline
88 & 99945 & 55 & 99566 & 434 & 428 & 49\tabularnewline
89 & 99976 & 24 & 99675 & 325 & 324 & 23\tabularnewline
90 & 99977 & 23 & 99756 & 244 & 243 & 22\tabularnewline
91 & 99982 & 18 & 99839 & 161 & 155 & 12\tabularnewline
92 & 99988 & 12 & 99865 & 135 & 135 & 12\tabularnewline
93 & 99994 & 6 & 99885 & 115 & 115 & 6\tabularnewline
94 & 99993 & 7 & 99911 & 89 & 88 & 6\tabularnewline
95 & 99998 & 2 & 99953 & 47 & 47 & 2\tabularnewline
96 & 99999 & 1 & 99965 & 35 & 35 & 1\tabularnewline
97 & 99999 & 1 & 99979 & 21 & 21 & 1\tabularnewline
98 & 100000 & 0 & 99973 & 27 & 27 & 0\tabularnewline
99 & 100000 & 0 & 99984 & 16 & 16 & 0\tabularnewline
100 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline

@@

\hypertarget{connectance}{\section{Sensitivity of connectance (C)
values}\label{connectance}}

### 2018-07-29 notebook/ms.Rmd
Discussion draft written and SI edits

Here I have shown that the stability of large systems might be often contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.

It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable. Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability. This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig. 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values. The mathematics underlying this shift in eigenvalue distribution have been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but have not been interpreted as rates of response of individual componetents to system perturbation.

Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems. The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.

I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks [@Allesina2012]. These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks). In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising [@Allesina2011]. When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information). Additionally, when each component of $\mathbf{M}$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).

Hence, managing the response rates of system components in a targeted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.

>>>>

Here I have shown that the stability of large systems might be often contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems. Variation in component response rate becomes more likely to be critical for stability as system size increases, and can ultimately increase the overall probability that system stability is observed above that predicted by May's[@May1972] classically derived $\sigma \sqrt{SC}$ criterion. The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.

It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable. Rather, systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig. 3). This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig. 1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues. The mathematics underlying this shift in eigenvalue distribution has been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual componetents to system perturbation.

The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability. The probability that some combination of component respose rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation. Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability. Hence, managing the response rates of system components in a targeted way could potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.

Interestingly, while complex systems were more likely to be stable given variation in component respose rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium. Feasibility is important to consider, particularly for the study of ecological networks of species[@Stone2017; @Grilli2017; @Dougoud2018] because population densities cannot realistically be negative. My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the stability of purely multi-species interaction networks (see also Supplementary Information). Nevertheless, ecological interactions do not exist in isolation in empirical systems[@Patel2018], but instead interact with evolutionary, abiotic, or social-economic systems. The relevance of component response rate for complex system stability should therefore not be ignored in the broader context of ecological communities.
